---
title: "DATA607: Week 1 Assignment"
author: "DH Kim"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

### Overview
The fivethirtyeight.com provides datasets for their analysis and storytelling. They have [a story on urbanization and election](https://fivethirtyeight.com/features/how-urban-or-rural-is-your-state-and-what-does-that-mean-for-the-2020-election/), great thinking linking the two interesting issues. Among two datasets provided for the article, I chose one for every census tract (named urbanization-census-tract.csv) to do some data work.

### Data Loading and Checking

```{r}
UrbanIndexRaw <- 
  read.csv(url("https://raw.githubusercontent.com/fivethirtyeight/data/master/urbanization-index/urbanization-census-tract.csv"))
dim(UrbanIndexRaw)
head(UrbanIndexRaw)
```
The raw data I imported has 8 variables for 73280 census tracts. Based on the resulting dataframe and [its details on their git repo]( https://github.com/fivethirtyeight/data/tree/master/urbanization-index), I change column names into a more meaningful way only for the selected columns below. Note that the dataset has urbanization data only, so election data is needed for future work. 

### Data Subsetting
```{r}
VariablesKeep <-
  c("statefips", "state", "lat_tract", "long_tract", 
    "population", "adj_radiuspop_5", "urbanindex")
VariablesName <-
  c("ID", "StateName", "Latitude", "Longitude",
    "Population", "UrbanPopulation", "UrbanIndex")
print(cbind(VariablesKeep,VariablesName))
```
```{r}
UrbanIndex <- UrbanIndexRaw[,VariablesKeep]
names(UrbanIndex) <- VariablesName
head(UrbanIndex)
rm(UrbanIndexRaw)
```
Compare it to the raw data. Now data is ready to generate your great stories.

### Findings and Conclusions
In the Week 1 Assignment, I take a close look at an urbanization dataset as a sample of data. It is great to know that data can be used to provide great stories. As shown in this Assignment, data work is very useful to do that. I think that This work may serve as a baseline work and can be extend for thinking and analysis. I need to be merged to an election data for the article. 
