---
title: "DATA607: Week 9 Assignment"
author: "DH Kim"
date: October 24, 2020
output: html_document
---
### Overview
In this Assignment, I obtain from [the New York Times Developer API website](https://developer.nytimes.com/apis) data on "most viewed articles" and "most shared articles" for the last seven days.

### Loading Packages
```{r load-packages, message=FALSE}
library(httr)
library(jsonlite)
library(tidyverse)
library(kableExtra)
library(DT)
```

### Most Shared Articles
The website provides API keys (See instructions in [Get Started](https://developer.nytimes.com/get-started) of the website) to get data. The GET() function in the httr library returns a response named *most_shared* whose content, as shown, is in JSON format. 
```{r}
nytd_most_shared_URL <- 
  "https://api.nytimes.com/svc/mostpopular/v2/shared/7/facebook.json?api-key=HUKJdm0PwgZd3c4mVStWWbIqbRObpRC8"
most_shared <- GET(nytd_most_shared_URL)
most_shared
http_status(most_shared)
#headers(most_shared)
```

The fromJSON() and data.frame() functions are used to convert the response into a dataframe. It has 22 columns for the 20 articles which most viewed for the last 7 days. Among other things, we can see article titles. 
```{r}
#most_shared$content
#content(most_shared, "raw")
#content(most_shared, "text")
most_shared_JSON <- fromJSON(content(most_shared, "text"))
most_shared_JSON$results$title
most_shared_df <- data.frame(most_shared_JSON$`results`)
glimpse(most_shared_df)
most_shared_df[,c("title")] %>%
  kbl() %>%
  kable_classic(full_width = F, html_font = "Cambria")
  

```

### Most Viewed Articles
Below is the most viewed articles from the last 7 days using the same API key.
```{r}
nytd_most_view_URL <- "https://api.nytimes.com/svc/mostpopular/v2/viewed/7.json?api-key=HUKJdm0PwgZd3c4mVStWWbIqbRObpRC8"
most_viewed <- GET(nytd_most_view_URL)
most_viewed_JSON <- fromJSON(content(most_viewed, "text"))
most_viewed_df <- data.frame(most_viewed_JSON$`results`)
most_viewed_df$title
```
### Analysis
Which sections are the most and the least viewed and shared?
```{r}
most_df <- rbind(most_shared_df, most_viewed_df)
most_df %>% 
  count(section) %>% 
  mutate (phat = n/sum(n)) %>% 
  arrange(desc(n))
ggplot(most_df, aes(x = section)) +
geom_bar() 
```

### Findings and Conclusions
I get data on articles from New York Times using API key that they provides. To do that, R provides the httr library and the GET() function in it. The most shared (in facebook) articles are most viewed articles are imported into data frame. Election-related articles are listed on top along with coronavirus-related articles. The U.S. section is the most viewed and shared and the technology and food sections are the least.   



