---
title: 'DATA607: Week 3 Assignment'
author: "DH Kim"
date: "September 12, 2020"
output:
  html_document:
    df_print: paged
  extra_dependencies:
  - geometry
  - multicol
  - multirow
  pdf_document: null
---

### Overview
This is string manipulation that we need to work with subset of string. It practices how to access characters and digits inside a string. We will deal with some useful functions and regular expressions. 

### Assignment 3.1

Provide code that identifies the majors that contain either "DATA" or "STATISTICS" by using the 173 majors listed in fivethirtyeight.com’s College Majors dataset.

The *grep()* function do the work.
```{r}
MajorData <- 
  read.csv(url("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"))
MajorInterested <- grep(pattern = "DATA|STATISTICS", MajorData$Major, value = TRUE)
length(MajorInterested)
print(MajorInterested)
```

### Assignment 3.2

Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

Step 1: Load the data into memory with string.

```{R}
rawData <- 
  "[1] \"bell pepper\"  \"bilberry\"     \"blackberry\"   \"blood orange\" \n
   [5] \"blueberry\"    \"cantaloupe\"   \"chili pepper\" \"cloudberry\" \n
   [9] \"elderberry\"   \"lime\"         \"lychee\"       \"mulberry\" \n
   [13] \"olive\"        \"salal berry\" "
writeLines(rawData)
```

Step 2: Extract only strings that we need.
```{R}
library(stringr)
inteData <- unlist(str_extract_all(rawData, "\"[[:alpha:] ]+\""))
writeLines(inteData)
```

Step 3: Add what we need and combine all together within a string.
```{r}
for (i in 1:length(inteData)){
  if (i==1) transData = str_c("c(",inteData[i],", ")
  if (i>=2 & i < length(inteData)) transData = str_c(transData,inteData[i],", ")
  if (i==length(inteData)) transData = str_c(transData, inteData[i],")")
}
writeLines(transData)
```

### Assignment 3.3

3.3.1
````
(.)\1\1
````
The regular expression ``"(.)\1\1"`` with double quotes matches any letter followed by ``\1\1``.  For comparison, ``"(.)\\1\\1"`` matches any letter which is repeated three times in a row.

```{r}
test <- c("777", "data\1\1", "anna", "2002", '"elle"')
str_view(test, "(.)\1\1", match = TRUE)
str_view(test, "(.)\\1\\1", match = TRUE)
```

3.3.2
````
"(.)(.)\\2\\1"
````
matches any two letters followed by the letters in reverse order like eppe (in pepper).

```{r}
str_view(test, "(.)(.)\\2\\1", match = TRUE)
```

3.3.3
````
(..)\1
````
``"(..)\1"`` matches any two letter followe by ``\1``.  ``"(..)\\1"`` matches any repeated pair of letters.

```{r}
str_view(test, "(..)\1", match = TRUE)
test <- c(test, "banana", "anan")
str_view(test, "(..)\\1", match = TRUE)
```

3.3.4
````
"(.).\\1.\\1"
````
matches two letters, followed by the same first letter (whatever the second letter is), followed by the first letter like anana or apaya.

```{r}
test <- c(test, "papaya")
str_view(test, "(.).\\1.\\1", match = TRUE)
```

3.3.5
````
"(.)(.)(.).*\\3\\2\\1"
````
matches from three letters to the same three letters in reverse order in any place.
```{r}
test <- c(test, "eeabcrtcbad")
str_view(test, "(.)(.)(.).*\\3\\2\\1", match = TRUE)
```

### Assignment 3.4
Construct regular expressions to match words that:

3.4.1

Start and end with the same character.
````
"^(.).*\\1$"
````
```{R}
str_view(test, "^(.).*\\1$", match = TRUE)
```


3.4.2

Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)

````
"(..).*\\1"
````
```{r}
test <- c(test, "church")
str_view(test, "(..).*\\1", match = TRUE)
```

3.4.3

Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

````
"(.).*\\1.*\\1"
````
```{r}
test <- c(test, "eleven")
str_view(test, "(.).*\\1.*\\1", match = TRUE)
```

### Findings and Conclusions
The best way to survive in the world of string manipulation and regular expression seems to be like this: practice, practice, and practice. In Project 1, we will use these string manipulation and regular expression to have a tidy data from a text file.

&nbsp;

&nbsp;

&nbsp;
