---
title: "DATA 607: Week 10 Assignment"
author: "DH Kim"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

### Introduction

In this Assignment, I conduct analysis of words in literature as a way to analyze sentiment in text using

- sentiment lexicons like afinn, bing, and nrc

- Jane Austin novels including Emma  

- wordcloud, analysis of the most common words

The analysis is extended with Loughran-McDonald Sentiment lexicon.

```{r load-packages, warning=FALSE, message=FALSE}
library(tidytext)
library(janeaustenr)
library(dplyr)
library(stringr)
library(tidyverse)
```

### Sentiment lexicons

Below are three different types of sentiment lexicons. Loughran-McDonald Sentiment lexicon is used later. In NRC, one word is mapped to multiple sentiment categories which are similar but different.

```{r code-chunk-label}
afinnSL <- get_sentiments("afinn")
table(afinnSL$value)
bingSL <- get_sentiments("bing")
table(bingSL$sentiment)
nrcSL <- get_sentiments("nrc")
table(nrcSL$sentiment)
```

### Sentiment joy in Emma

How does Emma express the sentiment of `joy`? We see in the book "Emma" top 10 words related to sentiment of `joy`.

```{r message=FALSE}
tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]", 
                                                 ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)
```

```{r message=FALSE}
nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")
tidy_Emma_joy <- tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)
tidy_Emma_joy
```
### Story Analysis

We see stories of novels in the perspective of sentiments of "positive" and "negative" which are defined in BING.

```{r message=FALSE}
library(tidyr)
library(ggplot2)

jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)

ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```

### Comparing lexicons
It compares three lexicons for story analysis of Pride and Prejudice with sentiment.
```{r message=FALSE}
pride_prejudice <- tidy_books %>%
  filter(book == "Pride & Prejudice")

afinn <- pride_prejudice %>%
  inner_join(get_sentiments("afinn")) %>%
  group_by(index = linenumber %/% 80) %>%
  summarise(sentiment = sum(value)) %>%
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(
  pride_prejudice %>%
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  pride_prejudice %>%
    inner_join(get_sentiments("nrc") %>%
      filter(sentiment %in% c(
        "positive",
        "negative"
      ))) %>%
    mutate(method = "NRC")
) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)

bind_rows(
  afinn,
  bing_and_nrc
) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y") +
  labs(title = "Sentiment Analysis of Pride and Prejudice")
```


### Wordcloud
```{r message=FALSE, warning=FALSE}
library(wordcloud)

tidy_books %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 80))
```

```{r message=FALSE}
library(reshape2)

tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(
    colors = c("blue", "red"),
    max.words = 80
  )
```

### Extension
Loughran-McDonald Sentiment lexicon is used for analysis.

```{r message=FALSE, warning=FALSE}
loughranSL <- get_sentiments("loughran")
table(loughranSL$sentiment)

tidy_books %>%
  inner_join(get_sentiments("loughran")) %>%
  filter(sentiment %in% c("positive","negative")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(
    colors = c("blue", "red"),
    max.words = 80
  )
```

### References

Silge, Julia, and David Robinson, *Text Mining with R: A Tidy Approach*, O'Reilly Media, 2020. https://www.tidytextmining.com/

