---
title: "Week 5 Assignment"
author: "DH Kim"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

### Overview
This Assignment imports, transforms, and analyzes data on the number of flights by arrival delays.

### Data Storage: .csv
We create data on arrival delays for two airlines (ALASKA and AMWEST) across five destinations (whose file is named *flightsInfo.csv*, as shown below, with wide data format) which is written by hand.
```{r load-packages, message=FALSE}
library(tidyverse)
library(RMySQL)
library(kableExtra)
```

```{r warning = FALSE}
numflightsTable <- read.csv("flightsInfo.csv", header=FALSE)
numflightsTable
```
### Data Storage: MySQL DB
The data can be stored in MySQL using the *flightsinfo.sql* (which is shown below) and *flightsInfo.csv*. We can access the database using the \textbf{RMySQL} library in R which will be shown soon.  
````
/*
  flightsinfo.sql
*/
DROP TABLE IF EXISTS flightsinfo;
CREATE TABLE flightsinfo
(
  airline varchar(50) NOT NULL,
  ontime varchar(25) NOT NULL,
  la int NULL,
  phoenix int NULL,
  sd int NULL,
  sf int NULL,
  seattle int NULL
);
SELECT * FROM flightsinfo;
LOAD DATA INFILE 'C:\\ProgramData\\MySQL\\MySQL Server 8.0\\Uploads\\flightsInfo.csv' 
INTO TABLE flightsinfo
FIELDS TERMINATED BY ',' 
ENCLOSED BY '"'
LINES TERMINATED BY '\n'
(airline, ontime, la, phoenix, sd, sf, seattle)
;
SELECT * FROM flightsinfo;
````

### MySQL in R
We connect to the MySQL DB in R and import data.
```{r}
flightsInfoDB <- dbConnect(MySQL(),
                           user='DATA607',
                           password='DATA607',
                           dbname='flightsinfo',
                           host='localhost')
dbListFields(flightsInfoDB, "flightsinfo")
queryResult <- dbSendQuery(flightsInfoDB, "SELECT * FROM flightsinfo")
numflightsTable <- fetch(queryResult,n=-1)
numflightsTable
```

### Tidying Data
The data is converted to a long format.
```{r}
numflights <- 
  pivot_longer(numflightsTable, cols = la:seattle,
    names_to = "destination", values_to = "count")
tibble(numflights)
```
Again, we convert it into a format that we will work with.
```{r}
numflights <- pivot_wider(numflights, names_from=ontime, values_from=count)
numflights
```
### Analyzing Data
Calculate and plot the proportion of on-time by destination and airline.
```{r warnings = FALSE}
numflights <- numflights %>%
  mutate(total = `on time` + delayed,
         onTime_ratio = round(`on time`/total, digits=2))
tibble(numflights)
```

```{r}
names(numflights) <- c("Airline", "Destination", "On-Time", "Delayed", "Total", "On-Time Rate")
numflights <- numflights[c("Airline", "Destination", 
                           "On-Time Rate", "On-Time",
                           "Delayed","Total")]
numflights %>%
     kbl() %>%
     kable_classic(full_width = F, html_font = "Cambria")
```

```{r, fig.width=6, fig.height=4}
ggplot(data = numflights, 
       aes(x = Destination, y = `On-Time Rate`, fill=Airline)) +
geom_bar(stat='identity',position=position_dodge()) +
scale_fill_brewer(palette="Paired")
 
```

### Findings and Conclusions
In this assignment, I deal with arrival delays data. I found out that the airport Phoenix has the highest on-time arrival rate by both airlines. In addition, ALASKA has higher on-time arrival rate than AMWEST. 