---
title: "DATA607: Week 2 Assignment"
author: "DH Kim"
date: "`r Sys.Date()`"
output:
    pdf_document:
        extra_dependencies: ["geometry", "multicol", "multirow"]
---

### Overview
This assignment exercises database as a method to store data and do data analysis. The week 1 assignment dealt with data with a single text file with extension csv. It is one way widely used for data analysis. Another is, as will be shown, relational database and sql which I in this assignment will explore with an example of movie recommendation. I think that, through the assignment, it is important to think about ways to store data, and the strength and weakness of the above two methods of data storage. 

### Movie Database
As will be seen, the survey results are stored in a database. Before going into that, I think it is worthy to note that. The table below shows data on movie ratings of six movies from five friends (A to  E), which is a good way to display data. The format is exactly what was in my head as a way to store the survey results in a csv file with comma delimited. From data perspective, it is called \textit{wide format}. Note that (1) there are multiple columns for movie ratings and (2) it can be converted to \textit{long format} which has just one column for the rating.  

\vspace{1em}

\begin{center}
\begin{tabular}{lccccc}
\hline
    & A & B & C & D & E\\
\hline
Kingsman &5&4&4&4&5\\
GoneGirl &4&3&4&4&2\\ 
Parasite &5&5&4&5&3\\ 
Robot\&Frank&4&5&4&5&5\\
CashBack&3&4&3&3&5\\
ARainyDayInNewYork&2&3&3&2&2\\
\hline
\end{tabular}
\end{center}

\vspace{1em}

The database used is Oracle database 11g Express Edition with SQL developer which can be freely downloaded from their websites. I think that it should be done with other databases, MySQL and PostgreSQL, and also R markdown which I am working with. The R markdown with DBI connection would be ideal but it remains as future work. Here, we will work with Oracle SQL developer with sql queries to create the movie database. The three steps do the work: (1) Create a user named \textit{moviedb}, (2) Create three tables named \textit{moving\_ratings}, \textit{friends}, and \textit{movies} and (3) insert data into relevant tables. It can be done by executing two sql files (\textit{setup\_moviedb.sql} and \textit{moviedb.sql}) attached. The more detailed procedure is as follows:

Step 1: To create a user, connect to the 'system' in Oracle SQL developer. It can be done with the green "+" icon. Then in the worksheet, execute \textit{setup\_moviedb.sql} file. Or type the "CREATE USER" command and press the green forward arrow to run it. 

Step 2: Connect to the \textit{moviedb} user created.

Step 3: To create tables and insert records, run \textit{moviedb.sql} using "CREATE TABLE" and "INSERT INTO" command.

### Movie Recommendation
We now have data on movie recommendation. SQL provides very powerful for data analysis (refer to \textit{prac.sql} file for some sql queries). The \textit{movies} and \textit{movie\_ratings} tables are shown in the Appendix. Note that, among other things, the rating data is in long format. 

Below shows a simple movie recommendation by just averaging movie ratings from friends. It is done by running the query: 

\vspace{1em}

```` 
SELECT movie_title, AVG(movie_rating) AS RecommendScore 
FROM movie_ratings
  JOIN movies
    ON movie_ratings.movie_id = movies.movie_id 
GROUP BY movie_title 
ORDER BY RecommendScore desc;
````

\vspace{1em}

Robot \& Frank (2013, Ford) is the highest.

\vspace{1em}

\begin{center}
\begin{tabular}{lc}
\hline
    &  RecommendScore\\
\hline
Robot\&Frank&4.6\\
Kingsman    &4.4\\
Parasite    &4.4\\ 
CashBack    &3.6\\
GoneGirl    &3.4\\ 
ARainyDayInNewYork&2.4\\
\hline
\end{tabular}
\end{center}


### Findings and Conclusions
During the database exercise with movie recommendation, I realized that it is interesting and important to understand various aspects of data storage and handling. It is not simple and straightforward to store data. So it needs to think about efficient ways of managing data.  

### Appendix
The \textit{movie\_ratings} table is as follows:
```{r fig.cap=c("movie_ratings table")}
knitr::include_graphics("pre_result2.pdf")
```
The \textit{movie} table is as follows:
```{r movie table}
knitr::include_graphics("pre_result1.pdf")
```

```{r}
knitr::include_graphics("result1.pdf")
knitr::include_graphics("result2.pdf")
```