---
title: "Project 3: The First Part"
author: "Cassie Boylan, DH Kim, Alexis Mekueko"
date: "`r Sys.Date()`"
header-includes: 
       - \usepackage{tikz}
       - \usetikzlibrary{positioning,shapes.multipart,shapes}
output: beamer_presentation
---
```{r setup, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

### Data Source and Loading
Packages used:
```{r load-packages, message=FALSE}
library(readxl)
library(httr)
```

E-Commerce Data:
```{r, size="tiny"}
retailURL <- 
    "http://archive.ics.uci.edu//ml//machine-learning-databases//00502//online_retail_II.xlsx"
```

Loading and Combining Data 
```{r, size = "tiny", results = "hide"}
GET(retailURL, write_disk(tempFileName <- tempfile(fileext = ".xlsx")))
```
```{r, size="tiny"}
retail_sheet_2009 <- read_excel(tempFileName, sheet = "Year 2009-2010")
retail_sheet_2010 <- read_excel(tempFileName, sheet = "Year 2010-2011")
retaildf <- rbind(retail_sheet_2009, retail_sheet_2010)
```


### Description of Data
```{r, size = "tiny"}
retaildf
```

### E-R Diagram

\tikzset{basic/.style={
        draw,
        rectangle split,
        rectangle split parts=2,
        rectangle split part fill={blue!20,white},
        minimum width=2.5cm,
        text width=2cm,
        align=left,
        font=\itshape
    },
    Diamond/.style={ diamond, 
                        draw, 
                        shape aspect=2, 
                        inner sep = 2pt,
                        text centered,
                        fill=blue!10!white,
                        font=\itshape
                      }}
\begin{tikzpicture}
\node[basic] (product) 
{product
\nodepart{second}
\underline{product\_id}\\
product\_name};
\node[basic,right=2cm of product] (invoice) 
{invoice
\nodepart{second}
\underline{invoice\_id}\\
customer\_id\\
invoice\_date};
\draw (product) -- (invoice) ;
\node[basic,right=2cm of invoice] (invoiceitem) 
{invoiceitem
\nodepart{second}
\underline{invoice\_id}\\
\underline{product\_id}\\
price\\
quantity};
\draw (invoice) -- (invoiceitem) ;
\node[basic,below=1.5cm of invoice] (customer) 
{customer
\nodepart{second}
\underline{customer\_id}\\
	country};
\draw (invoice) -- (customer) ;
\end{tikzpicture}

