---
title: 'E-Commerce Transaction Data: '
subtitle: "An Online Gift Shop Retailer Case"
author: "DATA607 Project 3 Presentation"
date: "Cassie Boylan, DH Kim, Alexis Mekueko"
output:
  beamer_presentation:
    theme: Antibes
    colortheme: dolphin
    toc: yes
    keep_tex: yes
    slide_level: 2
header-includes:
- \usepackage{tikz}
- \usetikzlibrary{positioning,shapes.multipart,shapes}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
---

```{r setup, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

## Who Presents What
\textbf{DH Kim}: Data Acquisition from Web

- Motivation, \textcolor{red}{Gift shop business}, \textcolor{red}{How to import Excel data from web}, Data structure

Cassie Boylan: Revenue Analysis from Invoices Data

- An Online gift shop retailer, Pre-processing data, Top 10 and bottom 10 gift items sold, Results of analysis

Alexis Mekueko: Data Management with Database

- Why database?: securing data, Theoretical E-R diagram, Normalizing: The customer table 

# Data Acquisition from Web
## Motivation

### Recent crisis of retail stores
### Linking business-related activities to real-world data 

\vspace{1em}

\begin{quote}
Virtually every aspect of business is now open to data collection and often instrumented for data collection: operations, manufacturing, \textbf{supply-chain management}, customer behaviour, market campaign performance, workflow procedures, and so on (page 1, Data Science for  Business)
\end{quote}

## Gift Shop Business and Transaction Data 

### An Online Retailer
- This is an UK-based online gift shop retailer selling gift goods to customers (mainly wholesalers) across countries. 
- Main items sold include Assorted Color Bird Wind Ornaments, Pink Cheery Lights, Floral Elephant Soft Toy, and so on.

### E-Commerce Invoice Data 
- Information on which items are sold, how many, and how much, who buys them, and when and what time are they ordered.  

- Data covering from 12/1/2009 to 12/9/2011, which is stored in an Excel file with two separate sheets.

## Importing Excel Data from the Web

Source: [\textcolor{brown}{The Website of UCI Machine Learning Repo} ](https://archive.ics.uci.edu/ml/datasets/Online+Retail+II)

Packages needed:
```{r load-packages, message=FALSE}
library(readxl)
library(httr)
```

The \textcolor{red}{GET()} and \textcolor{red}{read\_excel()} functions
```{r, size="tiny"}
retailURL <- 
    "http://archive.ics.uci.edu//ml//machine-learning-databases//00502//online_retail_II.xlsx"
```
```{r, size = "tiny", results = "hide"}
GET(retailURL, write_disk(tempFileName <- tempfile(fileext = ".xlsx")))
```
```{r, size="tiny"}
retail_sheet_2009 <- read_excel(tempFileName, sheet = "Year 2009-2010")
retail_sheet_2010 <- read_excel(tempFileName, sheet = "Year 2010-2011")
retaildf <- rbind(retail_sheet_2009, retail_sheet_2010)
```

## Description of Data 
### Invoices Data 
```{r message=FALSE,, size = "tiny"}
library(tidyverse)
glimpse(retaildf)
```

## Summary of Revenue and Transaction by Year
```{r message=FALSE, size = "tiny"}
library(stringr)
library(kableExtra)
retaildf$Date <- str_sub(retaildf$InvoiceDate, start=1, end=10)
retaildf$Year <- str_extract(retaildf$InvoiceDate, "\\d{4}")

summaReTr <- retaildf %>%
  filter(!is.na(Description)&!is.na(`Customer ID`)&Quantity > 0&Price > 0) %>%
  mutate(Revenue = round(Quantity*Price, digit = 2)) %>%
  group_by(Year) %>%
  summarise(Revenue = sum(Revenue),
            n_transactions = n_distinct(Invoice),
            n_obs = n(),
            First_Date = min(Date),
            Last_Date = max(Date))
kbl(summaReTr, booktabs = T) 
```

# Revenue Analysis from Invoices Data
## An Online Retailer Case

- This is Online Gift Shop Retailer

- Data from 12/1/2009 to 12/9/2011 daily including time information 

- The number of annual transactions is about 18,000 (Year 2010) 

- The annual revenue is about 8 millions (Year 2010)

## Pre-processing 1
```{r, message=FALSE, size="tiny"}
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyverse)
library(scales)
library(janitor)
library(epiDisplay)
```

```{r message=FALSE, size = "tiny"}
retaildf_2009 <- retaildf %>%
    filter(as.Date.POSIXct(InvoiceDate)=='2009-12-01') %>%
    filter(!is.na(Description) &!is.na(`Customer ID`) & Quantity > 0) %>%
    mutate(Dollar_Total = Quantity * Price) %>%
    group_by(StockCode, Description) %>%
    summarise(Total_Earned = sum(Dollar_Total),
              Total_Sold = sum(Quantity)) %>%
    arrange(desc(Total_Earned)) %>%
    ungroup() %>%
    mutate(Proportion_of_Revenue = scales::percent( Total_Earned/sum(Total_Earned)))
```

## Top 10 Gift Items Sold
```{r top-10, size = "tiny"}
top10 <- retaildf_2009 %>%
    filter(Total_Earned> 0) %>%
    slice_max(Total_Earned, n=10)
top10$StockCode <- NULL
kbl(top10, booktabs = T) %>%
kable_styling(latex_options = "striped")
```

## Bottom 10 Gift Items Sold
```{r bottom-10, size = "tiny"}
bottom10 <- retaildf_2009 %>%
    filter(Total_Earned > 0) %>%
    slice_min(Total_Earned, n=10)
bottom10$StockCode <- NULL
kbl(bottom10, booktabs = T) %>%
kable_styling(latex_options = "striped")
```

## Pre-processing 2
```{r, size = "tiny"}
proportion_totals <- retaildf_2009 %>%
    mutate(daily_revenue = case_when(
        between(Total_Earned, 1001, max(Total_Earned)) ~ "> $1000",
        between(Total_Earned, 500, 1000) ~ ">$500",
        between(Total_Earned, 301, 500) ~ ">$300",
        between(Total_Earned, 201, 300) ~">$200",
        between(Total_Earned, 101, 200) ~">$100",
        between(Total_Earned, 51, 100) ~">$50",
        between(Total_Earned, 25, 50) ~">=$25",
        between(Total_Earned, 0, 24) ~"<$25",
        TRUE ~ "loss")) %>%
    filter(daily_revenue!="loss")
```


## Results: Distribution of Revenue by SKU
```{r cumulative_proportion, , echo=FALSE, size = "tiny"}
tab1(proportion_totals$daily_revenue, sort.group = "increasing", cum.percent = TRUE, horiz=TRUE, main = "Distribution of Daily Revenue by SKU", cex=.7, cex.names = .8)
```

# Data Management with Database
## Why Database: An Example of Securing Data

```
# establishing the connectioon to SQL server to access db
con <- dbConnect(odbc(), 
  # server type
  Driver = "SQL Server", 
  #server name
  Server = "ATM\\ATMSERVER",
  # this is one of the db I want to import
  Database = "Data607_Project3_db", 
  UID = "Alex",
  # password required
  PWD = rstudioapi::askForPassword("Database password"), 
  ort = 1433)
}
```
\textcolor{red}{PWD = rstudioapi::askForPassword("Database password")}

## E-R Diagram
\tikzset{basic/.style={
        draw,
        rectangle split,
        rectangle split parts=2,
        rectangle split part fill={blue!20,white},
        minimum width=2.5cm,
        text width=2cm,
        align=left,
        font=\itshape
    },
    Diamond/.style={ diamond, 
                        draw, 
                        shape aspect=2, 
                        inner sep = 2pt,
                        text centered,
                        fill=blue!10!white,
                        font=\itshape
                      }}
\begin{tikzpicture}
\node[basic] (product) 
{product
\nodepart{second}
\underline{product\_id}\\
product\_name};
\node[basic,right=2cm of product] (invoice) 
{invoice
\nodepart{second}
\underline{invoice\_id}\\
customer\_id\\
invoice\_date};
\draw (product) -- (invoice) ;
\node[basic,right=2cm of invoice] (invoiceitem) 
{invoiceitem
\nodepart{second}
\underline{invoice\_id}\\
\underline{product\_id}\\
price\\
quantity};
\draw (invoice) -- (invoiceitem) ;
\node[basic,below=1.5cm of invoice] (customer) 
{customer
\nodepart{second}
\underline{customer\_id}\\
	country};
\draw (invoice) -- (customer) ;
\end{tikzpicture}

## Normalizing 
### A set of normalized tables
Customer, product, invoice, and invoiceitem tables

### An efficient way to storing data 
Easy for data maintenance and upgrade

### A secured way to access data
Control access to data with password

## An Example: The customer table
```{r}
varKeep <- c("Customer ID", "Country")
customerTable <- 
  retaildf[unique(retaildf$`Customer ID`), varKeep]
names(customerTable) <- c("CustomerID", "Country")
customerTable <- customerTable %>% 
  drop_na(CustomerID) %>%
  arrange(desc(CustomerID))
```


